Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization

### 2019-05-10 WRITEUP/ms.tex
merge
typos
up to sec 5

The virial theorem for a distribution of unmagnetized gas can be written as \citep{McKee92a}
\begin{equation}\label{eqn:virial_th_general}
\frac{1}{2}\ddot{\mathcal{I}} = 2(\mathcal{T} - \mathcal{T}_{\rm ext}) + \mathcal{W},
\end{equation}
where $\ddot{\mathcal{I}}$ is the second time derivative of the Lagrangian moment of inertia, $\mathcal{T}$ is the internal energy of the gas, $\mathcal{T}_{\rm ext}$ is external pressure support, and $\mathcal{W}$ is the gravitational energy.
%
Let us specialize to the case of a spherical self-gravitating molecular cloud of mass $M_{\rm gas}$, radius $R$, and root-mean-square velocity dispersion $\sigma$, accounting for both thermal and turbulent contributions.
%mm Neglecting magnetic field and
Defining $P_{\rm ext}$ as the external pressure, \Eq{virial_th_general} can be written as
\begin{equation}
\frac{1}{2}{\ddot{\mathcal{I}}} = 3 M_{\rm gas} \sigma^2 - 4\pi P_{\rm ext} R^3 - \Gamma\frac{GM_{\rm gas}^2}{R}\,,
\label{eqn:virial}
\end{equation}
where $\Gamma$ is a geometrical factor that is equal to 3/5 for a uniform sphere; in \Eq{virial} the terms on the right-hand side represent the
%mm  kinetic, external, and gravity pressure terms.
    kinetic energy, external pressure, and gravitational potential energy terms.

@@

In cases where stellar component plays an important role in the dynamics, the virial parameter becomes
\begin{equation}
\alpha_{\rm vir, tot} \equiv \frac{5 R}{G} \frac{M_{\rm gas} \sigma_{\rm gas}^2 + M_{\star} \sigma_\star^2}
					       {M_{\rm gas} + M_{\star} }\,,
\label{eqn:alpha_tot}
\end{equation}

\AP{shouldn t it be something like}
\begin{equation}
\alpha_{\rm vir, tot} \equiv \frac{5 R}{G (M_{\rm gas} + M_{\star})} \frac{M_{\rm gas} \sigma_{\rm gas}^2 + M_{\star} \sigma_\star^2}
					       {M_{\rm gas} + M_{\star} }\,,
\label{eqn:alpha_tot}
\end{equation}
\begin{equation}
\alpha_{\rm vir, tot} \equiv \frac{5 R \sigma_{\rm gas}^2}{G (M_{\rm gas} + M_{\star})}
\label{eqn:alpha_tot}
\end{equation}

where $M_\star$ is the stellar mass within the MCC volume.

@@

In \Fig{h2density}, we show the distribution of H$_2$ number density $n_{\rm H2}$ for \flower~ during its accreting phase, including the contribution from the gas within 3.5 kpc from the galaxy center.
%
We note that the distribution is almost flat for $n_{\rm H2}\gtrsim1$\,\cc and it samples the range of densities where clumps are found based on morphological analysis\footnote{For $n_{\rm H2}\gtrsim1$\,\cc, the fourth Minkowski functional of the H$_{2}$ density field is significantly larger than zero. This implies that the field is made of isolated components. See Fig. 6 in \citet{Pallottini17b}.} \citep{Pallottini17b}.
%
In each evolutionary stage, we identify MCCs by applying density cuts to the H$_2$ density distriution, i.e. $n_{\rm H2}\geq n_{\rm cut}$. We select 10 equally-spaced cuts in log scale: ${(n_{\rm cut}/{1 {\rm cm}^{-3})}}\eq[0.32, 0.53, 0.88, 1.45, 2.45, 4.08, 6.81, 11.36, 18.96, 31.62]$. We also tested variations in $n_{\rm cut}$ range and found no qualitative differences (see \Sec{ncut}). Note that with these choice we are including in the MCC also material that is not fully molecular ($n = 0.5 n_{\rm H2}$).

@@

The mass of an MCC is calculated from the uniformly-gridded 3D density field, integrating over the MCC volume $V$. The effective size is defined assuming spherical geometry, i.e. $R \equiv (3 V /4 \pi)^{1/3}$.
%
The full velocity dispersion of MCCs is calculated from the bulk velocity field ($\sigma_{\rm bulk}$), thermal sound speed ($c_s$), and the non-thermal velocity dispersion ($\sigma_{\rm NT}$)
\begin{equation}
\sigma_{\rm gas}^2 = \sigma_{\rm bulk}^2 + c_s^2 + \sigma_{\rm NT}^2.
\label{eqn:veldisp}
\end{equation}
%
In \obs of MCCs, the linewidth contribution of dense gas exceeds that from the diffuse gas. Therefore, when calculating global quantities for MCS, we always perform a mass-weighting. Since we operate on data on a unigrid, this is equivalent to density averaged quantities. In general, for the quantity $x$ in a MCC we write
\begin{equation}\label{eqn:defineaverage}
\langle x \rangle \equiv \frac{\sum_{i} \rho_i x_i }{\sum_i \rho_i}\,,
\end{equation}
where the sum is done for the cells indexed by $i$ composing the MCC. We can use the defintion \Eq{defineaverage} to write each term of the right-hand side of Equation~\ref{eqn:veldisp} as follows.
%
The bulk velocity dispersion is
\begin{equation}
\sigma_{\rm bulk}^2 = \frac{1}{{3}} \langle \left|\mathbf{v} - \langle \mathbf{v} \rangle  \right|^2\rangle \,.
\end{equation}
%
The thermal sound speed is calculated from the thermal pressure ($P_{\rm TH}$) through
\begin{equation}
c_s^2 = \frac{\sum_{i} \rho_i \left(P_{{\rm TH}, i}~{\rm k_B} / m_p / \rho_i \right)^2}{\sum_i \rho_i}\,,
c_s^2 = \left\langle \left( \frac{{\rm k_B} ~P_{\rm TH}}{{\rm m_p}\, n} \right)^2 \right\rangle\,,
\end{equation}
where the pressure is in units of K\,\cc, $\rm k_B$ is the Boltzmann constant, $m_p$ is the mass of a proton. Similarly, contribution from non-thermal energy is calculated from non-thermal pressure ($P_{\rm NT}$) as follows
\begin{equation}
\sigma_{\rm NT}^2 = \left\langle \left( \frac{{\rm k_B} ~P_{\rm NT}}{{\rm m_p}\, n} \right)^2 \right\rangle\,.
\end{equation}
%
Finally, the Mach number is related to the pressure terms as follows
\begin{equation}
\mathcal{M} = \left\langle \sqrt{ 1 + \frac{P_{\rm NT}}{P_{\rm TH}}} \right\rangle.
\label{eqn:MPress}
\end{equation}

>>>>

The virial theorem for a distribution of unmagnetized gas can be written as \citep{McKee92a}
\begin{equation}\label{eqn:virial_th_general}
\frac{1}{2}\ddot{\mathcal{I}} = 2(\mathcal{T} - \mathcal{T}_{\rm ext}) + \mathcal{W},
\end{equation}
where $\ddot{\mathcal{I}}$ is the second time derivative of the Lagrangian moment of inertia, $\mathcal{T}$ is the internal energy of the gas, $\mathcal{T}_{\rm ext}$ is external pressure support, and $\mathcal{W}$ is the gravitational energy.
%
Let us specialize to the case of a spherical self-gravitating molecular cloud of mass $M_{\rm gas}$, radius $R$, and root-mean-square velocity dispersion $\sigma$, accounting for both thermal and turbulent contributions.
%mm Neglecting magnetic field and
Defining $P_{\rm ext}$ as the external pressure, \Eq{virial_th_general} can be written as
\begin{equation}
\frac{1}{2}{\ddot{\mathcal{I}}} = 3 M \sigma^2 - 4\pi P_{\rm ext} R^3 - \Gamma\frac{GM^2}{R}\,,
\label{eqn:virial}
\end{equation}
where $\Gamma$ is a geometrical factor that is equal to 3/5 for a uniform sphere; in \Eq{virial} the terms on the right-hand side represent the
%mm  kinetic, external, and gravity pressure terms.
    kinetic energy, external pressure, and gravitational potential energy terms.

@@

In cases where stellar component plays an important role in the dynamics, the virial parameter becomes
\begin{equation}
\alpha_{\rm vir, tot} \equiv \frac{5 R}{G (M_{\rm gas} + M_{\star})} \frac{M_{\rm gas} \sigma_{\rm gas}^2 + M_{\star} \sigma_\star^2}
					       {M_{\rm gas} + M_{\star} }\,,
\label{eqn:alpha_tot}
\end{equation}
where $M_\star$ is the stellar mass within the MCC volume.
\DL{Should we leave out $\sigma_{\star}$ since stellar motions shouldn't affect the gas?}

@@

In \Fig{h2density}, we show the distribution of H$_2$ number density $n_{\rm H2}$ for \flower during its accreting phase, including the contribution from the gas within 3.5 kpc from the galaxy center.
%
We note that the distribution is almost flat for $n_{\rm H2}\gtrsim1$\,\cc and it samples the range of densities where clumps are found based on morphological analysis\footnote{For $n_{\rm H2}\gtrsim1$\,\cc, the fourth Minkowski functional of the H$_{2}$ density field is significantly larger than zero. This implies that the field is made of isolated components. See Fig. 6 in \citet{Pallottini17b}.} \citep{Pallottini17b}.
%
In each evolutionary stage, we identify MCCs by applying density cuts to the {\em H$_2$ density} distribution, i.e., MCCs 
are selected at $n_{\rm cut} \leq n_{\rm H2}$. We select 10 equally-spaced cuts in H$_2$ density in log scale: ${(n_{\rm cut}/{1 {\rm cm}^{-3})}}\eq[0.32, 0.53, 0.88, 1.45, 2.45, 4.08, 6.81, 11.36, 18.96, 31.62]$. 
% We also tested variations in $n_{\rm cut}$ range and found no qualitative differences (see \Sec{ncut}). 
Note that with these choice we are including MCCs that are not fully molecular ($n = 0.5 n_{\rm H2}$).

@@

The mass of an MCC is calculated from the uniformly-gridded 3D density field, integrating over the MCC volume $V$. The effective size is defined assuming spherical geometry, i.e. $R \equiv (3 V /4 \pi)^{1/3}$.
%
The full velocity dispersion of MCCs is calculated from the bulk velocity field ($\sigma_{\rm bulk}$), thermal sound speed ($c_s$), and the non-thermal velocity dispersion ($\sigma_{\rm NT}$)
\begin{equation}
\sigma_{\rm gas}^2 = \sigma_{\rm bulk}^2 + c_s^2 + \sigma_{\rm NT}^2.
\label{eqn:veldisp}
\end{equation}
%
In \obs of MCCs, the linewidth contribution of dense gas exceeds that from the diffuse gas. Therefore, when calculating global quantities for MCS, we always perform a mass-weighting. Since we operate on data on a unigrid, this is equivalent to density averaged quantities. In general, for the quantity $x$ in a MCC we write
\begin{equation}\label{eqn:defineaverage}
\langle x \rangle \equiv \frac{\sum_{i} \rho_i x_i }{\sum_i \rho_i}\,,
\end{equation}
where the sum is done for the cells indexed by $i$ composing the MCC. We can use the defintion \Eq{defineaverage} to write each term of the right-hand side of Equation~\ref{eqn:veldisp} as follows.
%
The bulk velocity dispersion is
\begin{equation}
\sigma_{\rm bulk}^2 = \frac{1}{{3}} \langle \left|\mathbf{v} - \langle \mathbf{v} \rangle  \right|^2\rangle \,.
\end{equation}
%
The thermal sound speed is calculated from the thermal pressure ($P_{\rm TH}$) through
\begin{equation}
c_s^2 = \left\langle \left( \frac{{\rm k_B} ~P_{\rm TH}}{{\rm m_p}\, n} \right)^2 \right\rangle\,,
\end{equation}
where the pressure is in units of K\,\cc, $\rm k_B$ is the Boltzmann constant, $m_p$ is the mass of a proton. Similarly, contribution from non-thermal energy is calculated from non-thermal pressure ($P_{\rm NT}$) as follows
\begin{equation}
\sigma_{\rm NT}^2 = \left\langle \left( \frac{{\rm k_B} ~P_{\rm NT}}{{\rm m_p}\, n} \right)^2 \right\rangle\,.
\end{equation}
%
Finally, the Mach number is related to the pressure terms as follows
\begin{equation}
\mathcal{M} = \left\langle \sqrt{ 1 + \frac{P_{\rm NT}}{P_{\rm TH}}} \right\rangle.
\label{eqn:MPress}
\end{equation}
